---
title: "Simulations"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# source functions
files.sources = paste0("../R/", list.files("../R"))
sapply(files.sources, source)
library(ggridges)
library(refund)

```


```{r}

# define components
K = 2; L = 2; t = sort(sample(1:100/100, 50, replace = FALSE)); p = length(t)

# define true W
spline = bs(t, knots = (0:6/6))

```


# Same Means Different PCs

### Define Setting

```{r}

# define Ws
coef11 = c(0,0,1,1,0,0,0,0,0,0)
temp = spline %*% coef11
pc11 = temp / as.numeric(sqrt(t(temp) %*% temp))

coef12 = c(0,0,0,0,1,1,0,0,0,0)
temp = spline %*% coef12
pc12 = temp / as.numeric(sqrt(t(temp) %*% temp))

coef13 = c(0,0,0,0,0,1,1,0,0,0)
temp = spline %*% coef13
pc13 = temp / as.numeric(sqrt(t(temp) %*% temp))

temp = rep(1,p)
pc14 = temp / as.numeric(sqrt(t(temp) %*% temp))

W.t = list()
W.t[[1]] = list(cbind(pc11, pc13), cbind(pc12, pc14))
W.t[[2]] = list(cbind(pc13, pc11), cbind(pc14, pc12))


# define true mu
mu.t = list()
mu.t[[1]] = cbind(rep(1,p), rep(1,p))
mu.t[[2]] = cbind(rep(1,p), rep(1,p))


# define true lambda
lambda.t = list()
lambda.t[[1]] = list(c(6,2), c(6,2))
lambda.t[[2]] = list(c(6,2), c(6,2))


# # simulate data
 sims = simulate_data(n = 50, m = 20, t = t,
             pi = c(0.5,0.5), rho = c(0.5, 0.5),
             mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

 sims$X %>%
   mutate(subj = factor(subj),
          trial_type = str_c("Condition ", trial_type),
          subjclus = str_c("Cluster ", subjclus)) %>%
   ggplot() +
   geom_line(aes(x = t, y = y, group = interaction(subj, trial_type, trial)), alpha = 0.2, size = 0.4) +
   facet_grid(cols = vars(trial_type), rows =  vars(subjclus)) +
   theme_bw() +
   theme(legend.position = "none",
         legend.title = element_blank()) +
   labs(x = "Time", y = "Simulated Activation") +
   guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))

ggsave("./Figures/simset1.png", width = 6, height = 4)

```


### Scenario 1: Exchangable clustering

```{r, eval = FALSE}

trialnum = c(10, 20, 40, 80); randlist = list()
for(j in 1:length(trialnum)) {

  for(i in 1:100) {

    # simulate data
    sims = simulate_data(n = 50, m = trialnum[j], t = t,
                pi = c(0.5,0.5), rho = c(0.5, 0.5),
                mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

    sm = sims$X %>%
      pivot_wider(id_cols = c(subj, trial, trial_type),
                  names_from = t, values_from = y, names_prefix = "t")

    matsm = sm %>%
      dplyr::select(-c(subj, trial, trial_type)) %>%
      as.matrix()

    fpca_sm = fpca.face(Y = matsm, pve = .95)

    clusex = sm %>%
      mutate(clusex = kmeans(fpca_sm$scores, 2)$cluster) %>%
      dplyr::select(subj, trial, trial_type, clusex) %>%
      group_by(subj) %>%
      summarise(m = mean(clusex)) %>%
      ungroup() %>%
      mutate(clus = ifelse(m <= 1.5, 1, 2)) %>%
      pull(clus)

    randlist[[i]] = tibble(iter = i, trialnum = trialnum[j], 
                           rand = rand.index(sims$clus, clusex),
                           scenario = 1)
    print(i)

  }
  randtib = do.call(rbind, randlist)
  write_csv(randtib, str_c("./Results/ex_s1_m", j, ".csv"))

}

exdat = rbind(read_csv("./Results/simslist_ex_s1_m1.csv"),
              read_csv("./Results/simslist_ex_s1_m2.csv"),
              read_csv("./Results/simslist_ex_s1_m3.csv")) 

```


### Scenario 1: Iterate Across Trial Numbers

```{r, eval = FALSE}

trialnum = c(10, 20, 40, 80); randlist = list()
for(j in 2:length(trialnum)) {

  simslist = list()
  for(i in 1:100) {

    # simulate data
    sims = simulate_data(n = 50, m = trialnum[j], t = t,
                pi = c(0.5,0.5), rho = c(0.5, 0.5),
                mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

    res = EM(sims$X, 2, 0.00001, cov.knots = 10, mu.knots = 10)

    randlist[[i]] = tibble(iter = i, trialnum = trialnum[j],
                           rand = rand.index(sims$clus, res$clus),
                           scenario = 1)
    print(i)

  }
  randtib = do.call(rbind, randlist)
  write_csv(randtib, str_c("./Results/s1_m", j, ".csv"))

}

```



### Scenario 1: Plot Results

```{r}

scen1 = rbind(rbind(read_csv("./Results/ex_s1_m1.csv"),
      read_csv("./Results/ex_s1_m2.csv"),
      read_csv("./Results/ex_s1_m3.csv"),
      read_csv("./Results/ex_s1_m4.csv")) %>%
    mutate(method = "Exchangeable"),

  rbind(read_csv("./Results/s1_m1.csv"),
      read_csv("./Results/s1_m2.csv"),
      read_csv("./Results/s1_m3.csv"),
      read_csv("./Results/s1_m4.csv")) %>%
    mutate(method = "FCM")
)

scen1 %>%
  ggplot() + 
  geom_density_ridges(aes(x = rand, y = method, fill = factor(trialnum)), alpha = 0.6) +
  theme_bw() + 
  labs(fill = "Number of Trials",
       x = "Rand Index") +
  theme(axis.title.y = element_blank(), 
        legend.position = "bottom")

```










## Same Means Difference in PCs in One Trial Type

### Define Scenario 

```{r}

# define Ws
coef11 = c(0,0,1,1,0,0,0,0,0,0)
temp = spline %*% coef11
pc11 = temp / as.numeric(sqrt(t(temp) %*% temp))

coef12 = c(0,0,0,0,1,1,0,0,0,0)
temp = spline %*% coef12
pc12 = temp / as.numeric(sqrt(t(temp) %*% temp))

coef13 = c(0,0,0,0,0,1,1,0,0,0)
temp = spline %*% coef13
pc13 = temp / as.numeric(sqrt(t(temp) %*% temp))

temp = rep(1,p)
pc14 = temp / as.numeric(sqrt(t(temp) %*% temp))

W.t = list()
W.t[[1]] = list(cbind(pc11, pc13), cbind(pc12, pc14))
W.t[[2]] = list(cbind(pc11, pc13), cbind(pc14, pc12))


# define true mu
mu.t = list()
mu.t[[1]] = cbind(rep(1,p), rep(1,p))
mu.t[[2]] = cbind(rep(1,p), rep(1,p))


# define true lambda
lambda.t = list()
lambda.t[[1]] = list(c(6,2), c(6,2))
lambda.t[[2]] = list(c(6,2), c(6,2))


# simulate data
sims = simulate_data(n = 50, m = 20, t = t,
            pi = c(0.5,0.5), rho = c(0.5, 0.5),
            mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

sims$X %>%
  mutate(subj = factor(subj),
         trial_type = str_c("Condition ", trial_type),
         subjclus = str_c("Cluster ", subjclus)) %>%
  ggplot() +
  geom_line(aes(x = t, y = y, group = interaction(subj, trial_type, trial)), alpha = 0.2, size = 0.4) +
  facet_grid(cols = vars(trial_type), rows =  vars(subjclus)) +
  theme_bw() +
  theme(legend.position = "none",
        legend.title = element_blank()) +
  labs(x = "Time", y = "Simulated Activation") +
  guides(color = guide_legend(override.aes = list(size = 3, alpha = 1)))

ggsave("./Figures/simset2.png", width = 6, height = 4)

```

## Scenario 2: Exchangeable Clustering

```{r, eval = FALSE}

trialnum = c(10, 20, 40, 80); randlist = list()
for(j in 1:length(trialnum)) {

  for(i in 1:100) {

    # simulate data
    sims = simulate_data(n = 50, m = trialnum[j], t = t,
                pi = c(0.5,0.5), rho = c(0.5, 0.5),
                mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

    sm = sims$X %>%
      pivot_wider(id_cols = c(subj, trial, trial_type),
                  names_from = t, values_from = y, names_prefix = "t")

    matsm = sm %>%
      dplyr::select(-c(subj, trial, trial_type)) %>%
      as.matrix()

    fpca_sm = fpca.face(Y = matsm, pve = .95)

    clusex = sm %>%
      mutate(clusex = kmeans(fpca_sm$scores, 2)$cluster) %>%
      dplyr::select(subj, trial, trial_type, clusex) %>%
      group_by(subj) %>%
      summarise(m = mean(clusex)) %>%
      ungroup() %>%
      mutate(clus = ifelse(m <= 1.5, 1, 2)) %>%
      pull(clus)

    randlist[[i]] = tibble(iter = i, trialnum = trialnum[j], 
                           rand = rand.index(sims$clus, clusex),
                           scenario = 2)
    print(i)

  }
  randtib = do.call(rbind, randlist)
  write_csv(randtib, str_c("./Results/ex_s2_m", j, ".csv"))

}

```



## Scenario 2: Iterate Across Trial Numbers

```{r, eval = FALSE}

trialnum = c(10, 20, 40, 80); randlist = list()
for(j in 4:length(trialnum)) {

  simslist = list()
  for(i in 1:100) {

    # simulate data
    sims = simulate_data(n = 50, m = trialnum[j], t = t,
                pi = c(0.5,0.5), rho = c(0.5, 0.5),
                mu = mu.t, W = W.t, lambda = lambda.t, sigma2 = 0.002)

    res = EM(sims$X, 2, 0.00001, cov.knots = 10, mu.knots = 10)

    randlist[[i]] = tibble(iter = i, trialnum = trialnum[j],
                           rand = rand.index(sims$clus, res$clus),
                           scenario = 2)
    print(i)

  }
  randtib = do.call(rbind, randlist)
  write_csv(randtib, str_c("./Results/s2_m", j, ".csv"))

}

```


## Plot All Scenarios

```{r}

scen2 = rbind(rbind(read_csv("./Results/ex_s2_m1.csv"),
      read_csv("./Results/ex_s2_m2.csv"),
      read_csv("./Results/ex_s2_m3.csv"),
      read_csv("./Results/ex_s2_m4.csv")) %>%
    mutate(method = "Exchangeable"),

  rbind(read_csv("./Results/s2_m1.csv"),
      read_csv("./Results/s2_m2.csv"),
      read_csv("./Results/s2_m3.csv"),
      read_csv("./Results/s2_m4.csv")) %>%
    mutate(method = "FCM")
)


# rbind(scen1, scen2) %>%
#   mutate(scenario = str_c("Scenario ", scenario)) %>%
#   ggplot() + 
#   geom_density_ridges(aes(x = rand, y = method, fill = factor(trialnum)), alpha = 0.7) +
#   theme_classic() + 
#   facet_wrap(~ scenario) +
#   labs(fill = "Number of Trials",
#        x = "Rand Index") +
#   theme(axis.title.y = element_blank(), 
#         legend.position = "bottom")


rbind(scen1, scen2) %>%
  mutate(scenario = str_c("Scenario ", scenario)) %>%
  mutate(method = ifelse(method == "Exchangeable", "Echangeable Clustering", "Unit-Level Clustering")) %>%
  ggplot() + 
  geom_boxplot(aes(x = method, y = rand, fill = factor(trialnum))) +
  theme_classic() + 
  facet_wrap(~ scenario) +
  labs(fill = "Number of Trials",
       y = "Rand Index") +
  theme(axis.title.x = element_blank(), 
        legend.position = "bottom")

ggsave("./Figures/res_boxplot.png", width = 7.5, height = 4)

```





```{r}

randdat %>%
  mutate(scenario = str_c("Scenario ", scenario)) %>%
  ggplot() + 
  geom_density_ridges(aes(x = rand, y = scenario, fill = factor(trialnum)), alpha = 0.6) +
  theme_bw() + 
  labs(fill = "Number of Trials",
       x = "Rand Index") +
  theme(axis.title.y = element_blank(), 
        legend.position = "bottom")

ggsave("./Figures/simres.png", width = 6.5, height = 5.5)


```






# rbind(tibble(pc = 1, clus = 1, type = 1, t = t, val = pc11),
#       tibble(pc = 2, clus = 1, type = 1, t = t, val = pc13),
#       tibble(pc = 1, clus = 1, type = 2, t = t, val = pc12),
#       tibble(pc = 2, clus = 1, type = 2, t = t, val = pc14),
#       tibble(pc = 1, clus = 2, type = 1, t = t, val = pc13),
#       tibble(pc = 2, clus = 2, type = 1, t = t, val = pc11),
#       tibble(pc = 1, clus = 2, type = 2, t = t, val = pc14),
#       tibble(pc = 2, clus = 2, type = 2, t = t, val = pc12)) %>%
#   mutate(clus = str_c("Cluster ", clus),
#          type = str_c("Scenario ", type),
#          pc = str_c("Component ", pc)) %>%
#   ggplot() + 
#   geom_line(aes(x = t, y = val, group = pc, color = pc), size = 0.8) +
#   facet_grid(cols = vars(clus), rows = vars(type)) +
#   theme_bw() +
#   theme(axis.title.y = element_blank(),
#         legend.position = "bottom",
#         legend.title = element_blank()) +
#   labs(x = "Time")
# 
# ggsave("./Figures/simpc1.png", width = 5.5, height = 4)

